---
title: "Lab Part 2"
author: "Laurel Abowd"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read_csv(here("data","lizards.csv"))
```

Get counts of lizards by species

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

The most commonly observed lizard here (UTST) is *Uta stansburiana*, or commonly the side-blotched lizard. We'll continue by exploring just that lizard. 

## Make a subset

Make a subset from `lizards` that only contains observations for: 
 
- side-blotched lizards (UTST)
- at sites "IBPE" (enclosure grassland site), "RABB" (rabbit mesquite site), and "EAST" (east tarbush site)
- where the tail condition is recorded as either "W" (whole) or "B" (broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("W", "B"))
```


## Find counts & proportions, put in a table

The `janitor::tabyl()` function gets this pretty close to contingency table format...

```{r}
tail_counts <- utst %>%
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
```

## Make contingency table and run chi-square

First, we see that the site is its own variable - when we run chi-square, we really only want the counts in the table. 

We can use `column_to_rownames` to convert entries in a column to stored rownames: 

```{r}
lizard_ct <- tail_counts %>%
  column_to_rownames(var = "site")
```

**Now** this is in shape to run chi-square. 

What are we asking? Are site and tail breakage independent? In other words: is there a significant association between site and tail breakage?

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2) #allows you to do inline referencing 
```

There is a significant association between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r round(lizard_tidy$statistic,2)`, *p* = `r round(lizard_tidy$p.value, 3)`). 

Remember: you'd also want to report the actual counts and proportions in each group! 

## END Part 2

